# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
1) We check for data directory, if it doesn't exist we create a data directory
2) We then download the file into the data directory
3) Unzip the file into data directory and read the file
4) We convert date field to Date type

```{r loadingData,echo=TRUE}
library(knitr)
if(!file.exists("./data")){
  dir.create("./data")
  fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(fileUrl,dest="./data/activity.zip")
  unzip("./data/activity.zip",exdir="./data")
  
  }
activity <- read.csv("./data/activity.csv")
activity$date <- as.Date(activity$date,"%Y-%m-%d")


```

## What is mean total number of steps taken per day?

```{r totalsteps,echo=TRUE}
head(activity)
library(reshape2)
activitymelt <- melt(activity,id=c("date","interval"),na.rm=TRUE)
activitydcast <- dcast(activitymelt,date~variable,sum,na.rm=TRUE)

mean(activitydcast$steps)

median(activitydcast$steps)


library(ggplot2)

ggplot(activitydcast, aes(x = steps)) + geom_histogram(binwidth=2000) + labs(x = "Number of Steps taken per day", 
        y = "Count") 

```


## What is the average daily activity pattern?

```{r}
activitymelt2 <- melt(activity,id=c("interval","date"),na.rm=TRUE)

activitydcast2 <- dcast(activitymelt2,interval~variable,mean)

head(activitydcast2,n=5)
summary(activitydcast2$steps)
activitydcast2$interval[which.max(activitydcast2$steps)]

ggplot(activitydcast2,aes(x=interval,y=steps))+geom_line()


```

## Imputing missing values

```{r}
head(is.na(activity))
head(activity[!complete.cases(activity),])
table(complete.cases(activity))
table(is.na(activity$steps))

missingindexes <- is.na(activity$steps)
missingactivity <- (activity[missingindexes,])
head(activitydcast2)

missing_indices <- which(is.na(activity$steps))
missing_steps <- unlist(lapply(missing_indices, FUN = function(idx) {
        interval = activity[idx, ]$interval
        activitydcast2[activitydcast2$interval == interval, ]$steps
    }))
new_steps <- activity$steps
new_steps[missing_indices] <- missing_steps
new_activity <- activity
new_activity$steps <- new_steps

head(activity)
head(new_activity)

new_activitymelt <- melt(new_activity,id=c("date","interval"),na.rm=TRUE)
new_activitydcast <- dcast(new_activitymelt,date~variable,sum,na.rm=TRUE)
mean(new_activitydcast$steps)
median(new_activitydcast$steps)

ggplot(new_activitydcast, aes(x = steps)) + geom_histogram(binwidth=2000) + labs(x = "Number of Steps taken per day", 
        y = "Count") 



```




## Are there differences in activity patterns between weekdays and weekends?

We do this comparison with the table with filled-in missing values.

1) Augment the table with a column that indicates the day of the week
2) Subset the table into two parts - weekends (Saturday and Sunday) and weekdays (Monday through Friday).
3) Tabulate the average steps per interval for each dataset.
4) Plot the two datasets side by side for comparison.

```{r,echo=TRUE}
str(activity)
new_activity$day <- as.factor(weekdays(new_activity$date))
weekend_activity <- subset(new_activity, day %in% c("Saturday", "Sunday"))
weekday_activity <- subset(new_activity, !day %in% c("Saturday", "Sunday"))

head(weekend_activity)
head(weekday_activity)
str(weekend_activity)

new <- data.frame(weekend_activity$steps,weekend_activity$interval,weekend_activity$date)
colnames(new) <- c("steps","interval","date")

#weekend_activity$steps <- as.integer(weekend_activity$steps)
#weekend_activity$interval <- as.numeric(weekend_activity$interval)

weekendmelt <- melt(new,id=c("interval","date"),na.rm=TRUE)

head(weekendmelt)
tail(weekendmelt)
str(weekendmelt)

weekendactivitydcast <- dcast(weekendmelt,interval~variable,mean,na.rm=TRUE)
head(weekendactivitydcast)

newweekday <- data.frame(weekday_activity$steps,weekday_activity$interval,weekday_activity$date)
colnames(newweekday) <- c("steps","interval","date")

weekdaymelt <- melt(newweekday,id=c("interval","date"),na.rm=TRUE)

head(weekdaymelt)
tail(weekdaymelt)
str(weekdaymelt)

weekdayactivitydcast <- dcast(weekdaymelt,interval~variable,mean,na.rm=TRUE)
head(weekdayactivitydcast)

weekdayactivitydcast$dayofweek <- "weekday"
weekendactivitydcast$dayofweek <- "weekend"



day_of_week_data <- rbind(weekdayactivitydcast, weekendactivitydcast)
day_of_week_data$dayofweek <- as.factor(day_of_week_data$dayofweek)

head(day_of_week_data)
tail(day_of_week_data)

plot_day_of_week_comparison <- function(dow_data) {
    ggplot(dow_data, aes(x = interval, y = steps)) + geom_line(color = "steelblue", 
        size = 1) + facet_wrap(~dayofweek, nrow = 2, ncol = 1) + labs(x = "Interval", 
        y = "Number of steps")
}

plot_day_of_week_comparison(day_of_week_data)

```
